---
title: "Crime Prediction"
author: "Jeyaraman Ramalingam"
date: "12/6/2021"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
    toc_float: yes
    theme: journal
---

## Libraries
```{r,eval=TRUE, message=FALSE,warning=FALSE, echo=F}
library(tidyverse)
library(caret)
library(pROC)
library(knitr)
library(Amelia)
library(naniar)
library(reshape2)
library(stats)
library(corrplot)
library(e1071)
library(jtools)
library(performance)
library(rJava)
library(glmulti)
library(cvms)
library(ROCR)
library(MASS)
library(rpart)
library(rpart.plot)
library(tree)
library(randomForest)
library(skimr)
library("PerformanceAnalytics")
library(caret)
library(dplyr)
library(gdata)
library(tidyverse)
library(Amelia)
library(mice)
library(kableExtra)
library(data.table)
library(e1071)
library(corrplot)
library(MASS)
library(caret)
library(tidyr)
library(dplyr)
library(data.table)
library(maps)
library("Hmisc")
library(kableExtra)
library(stringr)
library(ROCR)
library(DMwR2)
library(class)
library(ggplot2)
library(MLmetrics)
library(xgboost)
```

## Data Sources
```{r,eval=TRUE, message=FALSE,warning=FALSE, echo=F}

df_atlanta <- read.csv(file = 'datasets/Atlanta.csv')
df_atlanta <- df_atlanta[,c('Report.Number','Beat','Report.Date','UCR.Literal','Longitude','Latitude')]
df_atlanta <- df_atlanta[1:5000,]

factors <- factor(df_atlanta$UCR.Literal)
df_atlanta$OffenseCode <- as.numeric(factors)
df_atlanta<-within(df_atlanta,rm(UCR.Literal))


date = strptime(df_atlanta$Report.Date,"%Y-%m-%d")
df_atlanta["Year"] <- as.numeric(paste( format(date,"%Y")))
df_atlanta["Month"] <- as.numeric(paste(format(date,"%m")))
df_atlanta["Day"] <-  as.numeric(format(date,"%d"))

df_atlanta<-within(df_atlanta,rm(Report.Date))

df_atlanta<-na.omit(df_atlanta)


df_chicago <- read.csv(file = 'datasets/Chicago.csv')
df_chicago <- df_chicago[,c('ï..ID','Case.Number','Date','Primary.Type','Longitude','Latitude')]
df_chicago <- df_chicago[1:5000,]

factors <- factor(df_chicago$Primary.Type)
df_chicago$OffenseCode <- as.numeric(factors)
df_chicago<-within(df_chicago,rm(Primary.Type))


date = strptime(df_chicago$Date,"%m/%d/%Y %H:%M:%S")
df_chicago["Year"] <- as.numeric(paste( format(date,"%Y")))
df_chicago["Month"] <- as.numeric(paste(format(date,"%m")))
df_chicago["Day"] <-  as.numeric(format(date,"%d"))
df_chicago["Hour"] <-  as.numeric(format(date,"%H"))
df_chicago["HourC"] <-  as.numeric(format(date,"%H"))
df_chicago["DayDec"] <-  as.numeric(substr(format(date,"%d"),1,1))

df_chicago<-within(df_chicago,rm(Date))
df_chicago<-within(df_chicago,rm(Case.Number))

df_chicago<-na.omit(df_chicago)

df_la <- read.csv(file = 'datasets/LA.csv')
df_la <- df_la[,c('DR_NO','Date.Rptd','Crm.Cd.Desc','LON','LAT','AREA','Rpt.Dist.No')]
df_la <- df_la[1:5000,]

factors <- factor(df_la$Crm.Cd.Desc)
df_la$OffenseCode <- as.numeric(factors)
df_la<-within(df_la,rm(Crm.Cd.Desc))


date = strptime(df_la$Date.Rptd,"%m/%d/%Y %H:%M:%S")
df_la["Year"] <- as.numeric(paste( format(date,"%Y")))
df_la["Month"] <- as.numeric(paste(format(date,"%m")))
df_la["Day"] <-  as.numeric(format(date,"%d"))
df_la["Hour"] <-  as.numeric(format(date,"%H"))
df_la["HourC"] <-  as.numeric(format(date,"%H"))
df_la["DayDec"] <-  as.numeric(substr(format(date,"%d"),1,1))

df_la<-within(df_la,rm(Date.Rptd))
df_la<-within(df_la,rm(Year))
df_la<-within(df_la,rm(Hour))
df_la<-within(df_la,rm(HourC))

df_la<-na.omit(df_la)
df_ny <- read.csv(file = 'datasets/NY.csv')
df_ny <- df_ny[,c('CMPLNT_NUM','RPT_DT','OFNS_DESC','Longitude','Latitude','PD_CD','ADDR_PCT_CD')]
df_ny <- df_ny[1:5000,]

factors <- factor(df_ny$OFNS_DESC)
df_ny$OffenseCode <- as.numeric(factors)
df_ny<-within(df_ny,rm(OFNS_DESC))


date = strptime(df_ny$RPT_DT,"%m/%d/%Y")
df_ny["Year"] <- as.numeric(paste( format(date,"%Y")))
df_ny["Month"] <- as.numeric(paste(format(date,"%m")))
df_ny["Day"] <-  as.numeric(format(date,"%d"))
df_ny["DayDec"] <-  as.numeric(substr(format(date,"%d"),1,1))

df_ny<-within(df_ny,rm(RPT_DT))
df_ny<-within(df_ny,rm(Year))


df_ny<-na.omit(df_ny)


df_sfo <- read.csv(file = 'datasets/SFO.csv')
df_sfo <- df_sfo[,c('IncidntNum','Date','Category','Y','X','Incident.Code')]
df_sfo <- df_sfo[1:5000,]

factors <- factor(df_sfo$Category)
df_sfo$OffenseCode <- as.numeric(factors)
df_sfo<-within(df_sfo,rm(Category))


date = strptime(df_sfo$Date,"%m/%d/%Y")
df_sfo["Year"] <- as.numeric(paste( format(date,"%Y")))
df_sfo["Month"] <- as.numeric(paste(format(date,"%m")))
df_sfo["Day"] <-  as.numeric(format(date,"%d"))
df_sfo["Hour"] <-  as.numeric(format(date,"%H"))
df_sfo["HourC"] <-  as.numeric(format(date,"%H"))
df_sfo["DayDec"] <-  as.numeric(substr(format(date,"%d"),1,1))

df_sfo<-within(df_sfo,rm(Date))
df_sfo<-within(df_sfo,rm(Year))
df_sfo<-within(df_sfo,rm(Hour))
df_sfo<-within(df_sfo,rm(HourC))

df_sfo<-na.omit(df_sfo)


df_seattle <- read.csv(file = 'datasets/Seattle.csv')
df_seattle <- df_seattle[,c('Offense.ID','Report.DateTime','Offense','Longitude','Latitude','Beat')]
df_seattle <- df_seattle[1:5000,]

factors <- factor(df_seattle$Offense)
df_seattle$OffenseCode <- as.numeric(factors)
df_seattle<-within(df_seattle,rm(Offense))


date = strptime(df_seattle$Report.DateTime,"%m/%d/%Y %H:%M:%S")
df_seattle["Year"] <- as.numeric(paste( format(date,"%Y")))
df_seattle["Month"] <- as.numeric(paste(format(date,"%m")))
df_seattle["Day"] <-  as.numeric(format(date,"%d"))
df_seattle["Hour"] <-  as.numeric(format(date,"%H"))
df_seattle["HourC"] <-  as.numeric(format(date,"%H"))
df_seattle["DayDec"] <-  as.numeric(substr(format(date,"%d"),1,1))

df_seattle<-within(df_seattle,rm(Report.DateTime))
df_seattle<-within(df_seattle,rm(Beat))

names(df_atlanta)
names(df_chicago)
names(df_la)
names(df_ny)
names(df_sfo)
names(df_seattle)

head(df_atlanta)
head(df_chicago)
head(df_la)
head(df_ny)
head(df_sfo)
head(df_seattle)


```
## Input Statistics - Atlanta
```{r, message=FALSE,warning=FALSE, echo=F}

missmap(df_atlanta, main="Missing Values") 
colSums(is.na(df_atlanta))

#nonbinary <- c(1:10)
#X <- df_atlanta

#par(mfrow = c(3,4))
#for (i in nonbinary) {
#  hist(X[ ,i], xlab = names(X[i]), main = names(X[i]))
#  d <- density(X[,i])
#  plot(d, main = names(X[i]))
#  polygon(d, col="red")
#}


# Density plot to check normality
melt(df_atlanta, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_atlanta, skewness, function(x) skewness(x))

melt(df_atlanta, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_atlanta %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)


```

## Input Statistics - Chicago
```{r,eval=TRUE, message=FALSE,warning=FALSE, echo=F}
missmap(df_chicago, main="Missing Values") 
colSums(is.na(df_chicago))


# Density plot to check normality
melt(df_chicago, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_chicago, skewness, function(x) skewness(x))

melt(df_chicago, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_chicago %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)

```
## Input Statistics - Los Angeles
```{r, message=FALSE,warning=FALSE, echo=F}

missmap(df_la, main="Missing Values") 
colSums(is.na(df_la))

# Density plot to check normality
melt(df_la, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_la, skewness, function(x) skewness(x))

melt(df_la, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_la %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)

```
## Input Statistics - Seattle
```{r, message=FALSE,warning=FALSE, echo=F}
missmap(df_seattle, main="Missing Values") 
colSums(is.na(df_seattle))


# Density plot to check normality
melt(df_seattle, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_seattle, skewness, function(x) skewness(x))

melt(df_seattle, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_seattle %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)

```
## Input Statistics - San Francisco
```{r, message=FALSE,warning=FALSE, echo=F}

missmap(df_sfo, main="Missing Values") 
colSums(is.na(df_sfo))



# Density plot to check normality
melt(df_sfo, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_sfo, skewness, function(x) skewness(x))

melt(df_sfo, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_sfo %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)

```
## Input Statistics - New York
```{r, message=FALSE,warning=FALSE, echo=F}
missmap(df_ny, main="Missing Values") 
colSums(is.na(df_ny))

# Density plot to check normality
melt(df_ny, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=value))+geom_density(fill='gray')+facet_wrap(~variable, scales='free')+
  labs(title="Density Plot for Normality and Skewness") + 
  theme_classic()

# Skewness and outliers
#sapply(df_ny, skewness, function(x) skewness(x))

melt(df_ny, id.vars='OffenseCode') %>% mutate(OffenseCode = as.factor(OffenseCode)) %>% 
  ggplot(., aes(x=variable, y=value))+geom_boxplot(aes(fill=OffenseCode))+facet_wrap(~variable, dir='h',scales='free')+ labs(title="BoxPlot - Predictors Data Distribution with Target Variable")

df_ny %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", tl.cex=.8, diag = FALSE)

```



## Prediction Models - Atlanta
```{r, message=FALSE,warning=FALSE, echo=F}
indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]



train_x <- train[, -5]
train_y <- train[,5]

test_x <- test[, -5]
test_y <- test[,5]


dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model, data.frame(test_x))



mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()

dt_list <- data.frame('Decision Tree',mae,mse,rmse)

indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]



train_x <- train[, -4]
train_y <- train[,4]

test_x <- test[, -4]
test_y <- test[,4]


knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

knn_list <- data.frame('KNN',mae,mse,rmse)

indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]


model <-  naiveBayes(OffenseCode ~ ., data = train)
pred_y <- predict(model, test)
vector_final <- as.character(pred_y)
pred_y <- as.numeric(pred_y)
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)

indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]




train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]




classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

pred_y = predict(classifier, newdata = test_x)

cm = table(test_y, pred_y)
cm
postResample(test_y, pred_y)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)

indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]


model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")

mse = MSE(test$OffenseCode, round(pred))
mae = MAE(test$OffenseCode, round(pred))
rmse = RMSE(test$OffenseCode, round(pred))
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)

indexes = createDataPartition(df_atlanta$OffenseCode, p = .85, list = F)
train = df_atlanta[indexes, ]
test = df_atlanta[-indexes, ]




train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]


xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
atlanta_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)

```
## Prediction Models - Chicago
```{r, message=FALSE,warning=FALSE, echo=F}
indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]



train_x <- train[, -4]
train_y <- train[,4]

test_x <- test[, -4]
test_y <- test[,4]


dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model, data.frame(test_x))


mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()


dt_list <- data.frame('Decision Tree',mae,mse,rmse)

indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]



train_x <- train[, -4]
train_y <- train[,4]

test_x <- test[, -4]
test_y <- test[,4]


knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))



mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

knn_list <- data.frame('KNN',mae,mse,rmse)

indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]


model <-  naiveBayes(OffenseCode ~ ., data = train)
pred_y <- predict(model, test)
 
vector_final <- as.character(pred_y)
pred_y <- as.numeric(pred_y)

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)

indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]



train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]




classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

y_pred = predict(classifier, newdata = test_x)

cm = table(test_y, y_pred)
cm
postResample(test_y, y_pred)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)

indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]

model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")


mse = MSE(test$OffenseCode, round(pred))
mae = MAE(test$OffenseCode, round(pred))
rmse = RMSE(test$OffenseCode, round(pred))
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)

indexes = createDataPartition(df_chicago$OffenseCode, p = .85, list = F)
train = df_chicago[indexes, ]
test = df_chicago[-indexes, ]



train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]


xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)


mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
chicago_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)

```
## Prediction Models - Los Angeles
```{r, message=FALSE,warning=FALSE, echo=F}

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]

train_x <- train[, -6]
train_y <- train[,6]

test_x <- test[, -6]
test_y <- test[,6]



dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model.pruned, data.frame(test_x))


mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()


dt_list <- data.frame('Decision Tree',mae,mse,rmse)

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]

train_x <- train[, -6]
train_y <- train[,6]

test_x <- test[, -6]
test_y <- test[,6]


knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))



mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

knn_list <- data.frame('KNN',mae,mse,rmse)

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]


model <-  naiveBayes(OffenseCode ~ ., data = train)
predictions <- predict(model, test)
vector_final <- as.character(predictions)

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]

train_x <- data.matrix(train[, -6])
train_y <- train[,6]

test_x <- data.matrix(test[, -6])
test_y <- test[,6]

classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

pred_y = predict(classifier, newdata = test_x)

cm = table(test_y, pred_y)
cm
postResample(test_y, pred_y)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]

model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")

mse = MSE(test$OffenseCode, pred)
mae = MAE(test$OffenseCode, pred)
rmse = RMSE(test$OffenseCode, pred)
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)

indexes = createDataPartition(df_la$OffenseCode, p = .85, list = F)
train = df_la[indexes, ]
test = df_la[-indexes, ]

train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]




xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
la_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)


```
## Prediction Models - Seattle
```{r, message=FALSE,warning=FALSE, echo=F}

indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


train_x <- train[, -4]
train_y <- train[,4]

test_x <- test[, -4]
test_y <- test[,4]



dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model, data.frame(test_x))



mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()

dt_list <- data.frame('Decision Tree',mae,mse,rmse)


indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


train_x <- train[, -4]
train_y <- train[,4]

test_x <- test[, -4]
test_y <- test[,4]


knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))



mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

knn_list <- data.frame('KNN',mae,mse,rmse)


indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


model <-  naiveBayes(OffenseCode ~ ., data = train)
pred_y <- predict(model, test)
 
vector_final <- as.character(pred_y)
pred_y <- as.numeric(pred_y)
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)


indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]


classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

y_pred = predict(classifier, newdata = test_x)

cm = table(test_y, y_pred)
cm
postResample(test_y, y_pred)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)


indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")

mse = MSE(test$OffenseCode, round(pred))
mae = MAE(test$OffenseCode, round(pred))
rmse = RMSE(test$OffenseCode, round(pred))
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)


indexes = createDataPartition(df_seattle$OffenseCode, p = .85, list = F)
train = df_seattle[indexes, ]
test = df_seattle[-indexes, ]


train_x <- data.matrix(train[, -4])
train_y <- train[,4]

test_x <- data.matrix(test[, -4])
test_y <- test[,4]


xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
seattle_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)



```
## Prediction Models - San Francisco
```{r, message=FALSE,warning=FALSE, echo=F}

indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]

train_x <- train[, -5]
train_y <- train[,5]

test_x <- test[, -5]
test_y <- test[,5]


dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model, data.frame(test_x))


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()

dt_list <- data.frame('Decision Tree',mae,mse,rmse)
indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]

train_x <- train[, -5]
train_y <- train[,5]

test_x <- test[, -5]
test_y <- test[,5]


knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))




mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


knn_list <- data.frame('KNN',mae,mse,rmse)

indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]



model <-  naiveBayes(OffenseCode ~ ., data = train)
pred_y <- predict(model, test)
 
vector_final <- as.character(pred_y)
pred_y <- as.numeric(pred_y)
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)

indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]

train_x <- data.matrix(train[, -5])
train_y <- train[,5]

test_x <- data.matrix(test[, -5])
test_y <- test[,5]




classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

y_pred = predict(classifier, newdata = test_x)

cm = table(test_y, y_pred)
cm
postResample(test_y, y_pred)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)

indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]


model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")

mse = MSE(test$OffenseCode, round(pred))
mae = MAE(test$OffenseCode, round(pred))
rmse = RMSE(test$OffenseCode, round(pred))
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)

indexes = createDataPartition(df_sfo$OffenseCode, p = .85, list = F)
train = df_sfo[indexes, ]
test = df_sfo[-indexes, ]

train_x <- data.matrix(train[, -5])
train_y <- train[,5]

test_x <- data.matrix(test[, -5])
test_y <- test[,5]




xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)

mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
sfo_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)


```
## Prediction Models - New York
```{r, message=FALSE,warning=FALSE, echo=F}

indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]

train_x <- train[, -6]
train_y <- train[,6]

test_x <- test[, -6]
test_y <- test[,6]

dt_model<-rpart(train_y~., data = data.frame(train_x,train_y), control = rpart.control(cp = 0.00001))



dt_model.pruned = prune(dt_model, cp = 0.0001)

plot(dt_model.pruned)

text(dt_model.pruned, cex = 0.9, xpd = TRUE)

pred_y = predict(dt_model, data.frame(test_x))



mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


x = 1:length(test_y)

plot(x, test_y, col = "red", type = "l", lwd=2,
     main = "Boston housing test data prediction")
lines(x, pred_y, col = "blue", lwd=2)
legend("topright",  legend = c("original-medv", "predicted-medv"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid()


dt_list <- data.frame('Decision Tree',mae,mse,rmse)
indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]

train_x <- train[, -6]
train_y <- train[,6]

test_x <- test[, -6]
test_y <- test[,6]



knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))




mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)


knn_list <- data.frame('KNN',mae,mse,rmse)

indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]



model <-  naiveBayes(OffenseCode ~ ., data = train)
pred_y <- predict(model, test)
 
vector_final <- as.character(pred_y)
pred_y <- as.numeric(pred_y)
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

nb_list <- data.frame('Naive Bayes',mae,mse,rmse)

indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]

train_x <- data.matrix(train[, -6])
train_y <- train[,6]

test_x <- data.matrix(test[, -6])
test_y <- test[,6]


classifier = randomForest(x = train_x,y = train_y,ntree = 500, random_state = 0)

y_pred = predict(classifier, newdata = test_x)

cm = table(test_y, y_pred)
cm
postResample(test_y, y_pred)
varImpPlot(classifier)

mtry = sqrt(30)

importance = importance(classifier)
importance
mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

rf_list <- data.frame('Random Forest',mae,mse,rmse)

indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]

model_reg = svm(OffenseCode~., data=train)


pred = predict(model_reg, test)

x = 1:length(test$OffenseCode)
plot(x, test$OffenseCode, pch=18, col="red")
lines(x, pred, lwd="1", col="blue")

mse = MSE(test$OffenseCode, round(pred))
mae = MAE(test$OffenseCode, round(pred))
rmse = RMSE(test$OffenseCode, round(pred))
r2 = R2(test$OffenseCode, pred, form = "traditional")

cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)


svm_list <- data.frame('SVM',mae,mse,rmse)

indexes = createDataPartition(df_ny$OffenseCode, p = .85, list = F)
train = df_ny[indexes, ]
test = df_ny[-indexes, ]

train_x <- data.matrix(train[, -6])
train_y <- train[,6]

test_x <- data.matrix(test[, -6])
test_y <- test[,6]


xgb_train = xgb.DMatrix(data = train_x, label = train_y)
xgb_test = xgb.DMatrix(data = test_x, label = test_y)

xgbc = xgboost(data = xgb_train, max.depth = 2, nrounds = 50)

pred_y = predict(xgbc, xgb_test)


mse = mean((test_y - round(pred_y))^2)
mae = caret::MAE(test_y, round(pred_y))
rmse = caret::RMSE(test_y, round(pred_y))


cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)

xg_list <- data.frame('xGBoost',mae,mse,rmse)

colnames(dt_list)[1] <- "Model_Name"
colnames(knn_list)[1] <- "Model_Name"
colnames(nb_list)[1] <- "Model_Name"
colnames(rf_list)[1] <- "Model_Name"
colnames(svm_list)[1] <- "Model_Name"
colnames(xg_list)[1] <- "Model_Name"
ny_metrics <- rbind(dt_list,knn_list,nb_list,rf_list,svm_list,xg_list)

```

## Model Performance
```{r, message=FALSE,warning=FALSE, echo=F}


atlanta_metrics$City <- 'Atlanta'
seattle_metrics$City <- 'Seattle'
sfo_metrics$City <- 'San Francisco'
ny_metrics$City <- 'New York'
la_metrics$City <- 'Los Angeles'
chicago_metrics$City <- 'Chicago'

Model_Metrics <- rbind(atlanta_metrics,seattle_metrics,sfo_metrics,ny_metrics,la_metrics,chicago_metrics)



```


## ARIMA Model Forecasting - Atlanta
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)
library(dplyr)

df_atlanta <- read.csv(file = 'datasets/Atlanta/atlanta_final.csv')
df_atlanta_summary<-df_atlanta %>%
  group_by(Rpt_year) %>%
  summarise(n=n()) 

#convert to time series
tsData <- ts(df_atlanta_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)

#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))
qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Atlanta Crimes") 

```
## ARIMA Model Forecasting - Chicago
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)

#import data
df_chicago <- read.csv(file = 'datasets/Chicago/chicago_final.csv')
date = strptime(df_chicago$Date,"%m/%d/%Y %H:%M:%S")
df_chicago["Year"] <- as.numeric(paste( format(date,"%Y")))

df_chicago_summary<-df_chicago %>%
  group_by(Year) %>%
  summarise(n=n()) 

#convert to time series
tsData <- ts(df_chicago_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)


#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Chicago Crimes") 

```
## ARIMA Model Forecasting - Los Angeles
```{r, message=FALSE,warning=FALSE, echo=F}

#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)

#import data
df_la <- read.csv(file = 'datasets/Los Angeles/la_final.csv')
date = strptime(df_la$Date.Rptd,"%m/%d/%Y %H:%M:%S")
df_la["Year"] <- as.numeric(paste( format(date,"%Y")))
df_la_summary<-df_la %>%
  group_by(Year) %>%
  summarise(n=n()) 

#convert to time series
tsData <- ts(df_la_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)


#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Chicago Crimes") 


```
## ARIMA Model Forecasting - Seattle
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)

#import data
df_seattle <- read.csv(file = 'datasets/Seattle/seattle_final.csv')
date = strptime(df_seattle$Report.DateTime,"%m/%d/%Y %H:%M:%S")
df_seattle["Year"] <- as.numeric(paste( format(date,"%Y")))
df_seattle_summary<-df_seattle %>%
  group_by(Year) %>%
  summarise(n=n()) 


#convert to time series
tsData <- ts(df_seattle_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)


#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Chicago Crimes") 

```
## ARIMA Model Forecasting - San Francisco
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)

#import data
df_sfo <- read.csv(file = 'datasets/San Francisco/2009.csv')
date = strptime(df_sfo$Date,"%m/%d/%Y")
df_sfo["Year"] <- as.numeric(paste( format(date,"%Y")))

df_sfo_summary<-df_sfo %>%
  group_by(Year) %>%
  summarise(n=n()) 


#convert to time series
tsData <- ts(df_sfo_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)

#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(2,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Chicago Crimes") 

```
## ARIMA Model Forecasting - New York
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)

#import data
df_ny <- read.csv(file = 'datasets/New York/ny_final.csv')
date = strptime(df_ny$CMPLNT_FR_DT,"%m/%d/%Y")
df_ny["Year"] <- as.numeric(paste( format(date,"%Y")))
df_ny_summary<-df_ny %>%
  group_by(Year) %>%
  summarise(n=n()) 


#convert to time series
tsData <- ts(df_ny_summary[2],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)

#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for Chicago Crimes") 

```

## Overall USA Crime
```{r, message=FALSE,warning=FALSE, echo=F}
#Load Libraries
library("fUnitRoots")
library(ggplot2)
library(zoo)
library(lmtest)
library("forecast")
library(FitAR)
library(tidyverse)
library(dplyr)
#import data
df_seattle <- read.csv(file = 'datasets/Seattle/seattle_final.csv')
date = strptime(df_seattle$Report.DateTime,"%m/%d/%Y %H:%M:%S")
df_seattle["Year"] <- as.numeric(paste( format(date,"%Y")))
df_seattle_summary<-df_seattle %>%
  group_by(Year) %>%
  summarise(n=n()) 

df_atlanta_summary$Year <- df_atlanta_summary$Rpt_year
df_atlanta_summary<-within(df_atlanta_summary,rm(Rpt_year))
df_us_summary <- rbind(df_atlanta_summary,df_chicago_summary,df_seattle_summary,df_sfo_summary,df_ny_summary,df_la_summary)

df_us_summary_cnt<-df_us_summary %>%
  dplyr::filter(Year>=2000) %>%
  group_by(Year) %>%
  summarise(n=sum(n)) 

  #convert to time series
tsData <- ts(df_us_summary[1],start=2009,end=2020, frequency = 1)
plot(tsData)

#detemine stationarity of data
urkpssTest(tsData, type = c("tau"), lags = c("short"),use.lag = NULL, doplot = TRUE)
tsstationary<-diff(tsData, differences=1)
plot(tsstationary)
acf(tsData,lag.max=34)

#fit the model
fitARIMA<-arima(tsData, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
fitARIMA
#significance of coefficients
coeftest(fitARIMA)
par(mfrow=c(1,1))
acf(fitARIMA$residuals)

par(mfrow=c(2,1))

qqnorm(fitARIMA$residuals)
qqline(fitARIMA$residuals)

auto.arima(tsData, trace=TRUE)

#forcast future values
par(mfrow=c(1,1))
predict(fitARIMA,n.ahead = 5)
futurVal <- forecast(fitARIMA,h=5)
plot(futurVal,xlab="Year",ylab="Number of Crimes",main="ARIMA Forecast for USA Crimes",ylim = c(10000, 400000)) 

```
